Analysis and figures

### 2025-02-03 Analysis.qmd
Completed first-pass revisions for Ecography

@@

@@

```{r Objective proportions}
stat_objectiveProportions <- sheet_Objectives |>
  mutate(TotalArticles = length(unique(`Review ID`))) |>
  pivot_longer(c("Demonstrate methods?", "Estimate trends?", 
                 "Identify drivers?", "Test relationships?",
                 "Predict spatially?", "Predict temporally?"),
               names_to = "Objective", values_to = "ObjectiveIncluded") |>
  filter(ObjectiveIncluded == "YES") |>
  summarise(ObjArticles = n(), 
            .by = c(Objective, TotalArticles)) |>
  mutate(PropArticles = ObjArticles/TotalArticles)
stat_objectiveProportions
```

@@

>>>>

## Authorship

```{r Authorship_categories}
sheet_Objectives |>
  separate_longer_delim(`Authorship categories`, delim = ", ") |>
  mutate(n_Overall = length(unique(`Review ID`))) |>
  summarise(PropAuthor = (n()/mean(n_Overall))*100,
            .by = `Authorship categories`)

sheet_Objectives |>
  separate_longer_delim(`Authorship categories`, delim = ", ") |>
  summarise(nPer = n(),
            .by = `Review ID`) |>
  summarise(AffilCount = n(),
            .by = nPer) (92-34)/92

sheet_ArticleData |>
  separate_longer_delim(`Countries of affiliation`, delim = ", ") |>
  summarise(nCountry = n(),
            .by = `Countries of affiliation`) 
```

@@

## Covariate summaries

#### Quantity

```{r Covariate_Quantity}
covariateFreq <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, Covariate, Category, n_overall) |>
  filter(!is.na(Covariate)) %>%
  rbind(mutate(., Covariate = Category)) %>%
  rbind(mutate(., Covariate = "ANY", Category = "ANY")) |>
  distinct() |>
  summarise(PctCategory = (n()/mean(n_overall))*100,
            .by = c(Covariate, Category)) |>
  mutate(Included = TRUE) %>%
  rbind(mutate(., PctCategory = 100-PctCategory, Included = FALSE)) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species"))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>

  ggplot(aes(x = Covariate, y = PctCategory, fill = Included)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c(TRUE, FALSE), 
                    values = c("goldenrod", "gray50")) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(Category), scales = "free_x", space = "free_x") +
  labs(y = "Percent\nof models", x = "") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(family = "Helvetica", 
                                   angle = 90, vjust = 0.5, hjust = 1),
        strip.text.x = element_text(family = "Helvetica"))
covariateFreq
```

#### Dynamism

```{r Covariate_Dynamism}
covariateDyn <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, `Parameter type`, Covariate, Category, Variation, n_overall) |>
  filter(!is.na(Covariate)) |>
  filter(`Parameter type` != "Initial occupancy") %>%
  rbind(rbind(mutate(., Covariate = Category),
              mutate(., Covariate = "ANY", Category = "ANY"))) |>
  mutate(n_Covariate = n(),
         .by = Covariate) |>
  mutate(Dynamism = case_when(Variation %in% c("SISE", "SU", "SE") ~ 1,
                              .default = 0)) |>
  summarise(PctDyn = (sum(Dynamism)/mean(n_Covariate))*100,
            .by = c(Covariate, Category)) |>
  mutate(Dyn = TRUE) %>%
  rbind(mutate(., PctDyn = 100-PctDyn, Dyn = FALSE)) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species"))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>

  ggplot(aes(x = Covariate, y = PctDyn, fill = Dyn)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c(TRUE, FALSE), 
                    values = c("aquamarine3", "gray50")) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(Category), scales = "free_x", space = "free_x") +
  labs(y = "Percent\ndynamic", x = "") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(family = "Helvetica", 
                                   angle = 90, vjust = 0.5, hjust = 1),
        strip.text.x = element_blank(), strip.background = element_blank())
covariateDyn
```

#### Observation

```{r Covariate_Observed}
covariateObs <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, Covariate, Category, Observation, n_overall) |>
  filter(!is.na(Covariate)) %>%
  rbind(rbind(mutate(., Covariate = Category),
              mutate(., Covariate = "ANY", Category = "ANY"))) |>
  mutate(n_Covariate = n(),
         .by = Covariate) |>
  mutate(Observed = case_when(Observation == "D" ~ 1,
                              .default = 0)) |>
  summarise(PctObs = (sum(Observed)/mean(n_Covariate))*100,
            .by = c(Covariate, Category)) |>
  mutate(Obs = TRUE) %>%
  rbind(mutate(., PctObs = 100-PctObs, Obs = FALSE)) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species"))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>

  ggplot(aes(x = Covariate, y = PctObs, fill = Obs)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c(TRUE, FALSE), 
                    values = c("cadetblue3", "gray50")) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(Category), scales = "free_x", space = "free_x") +
  labs(y = "Percent directly\nobserved", x = "") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(family = "Helvetica", 
                                   angle = 90, vjust = 0.5, hjust = 1),
        strip.text.x = element_blank(), strip.background = element_blank())
covariateObs

#### Complexity

```{r Covariate_Complexity}
covariateComplex <- sheet_Covariates |>
  separate_longer_delim(Covariates, delim = ", ") |>
  separate_wider_delim(Covariates, delim = "_", 
                       names = c("Covariate", "Variation", 
                                 "Observation", "Relation"),
                       too_few = "align_start") |>
  separate_wider_delim(Covariate, delim = "-",
                       names = c("Category", "Covariate"),
                       too_few = "align_start") |>
  # Merge some covariate categories
  mutate(Covariate = case_when(Covariate == "OTHER" & Category == "E" ~ "OTHER-E",
                               Covariate == "OTHER" & Category == "S" ~ "OTHER-S",
                               Covariate %in% c("HABT", "HABA", "LACO") ~ "Habitat",
                               Covariate %in% c("CLWE", "DISA", "DIST") ~ "Climate",
                               Covariate %in% c("SPAT", "GEOM", "CONN") ~ "Spatial",
                               Covariate %in% c("ANTH", "AGRI", "PROT") ~ "Human",
                               .default = Covariate)) |>
  # Combine Review ID & Model ID, and calculate the number of studies and number of studies with each parameter
  mutate(studyID = paste(`Review ID`, `Model ID`), sep = "_") |>
  mutate(n_overall = length(unique(studyID))) |>
  select(studyID, Covariate, Category, Relation, n_overall) |>
  filter(!is.na(Covariate)) %>%
  rbind(rbind(mutate(., Covariate = Category),
              mutate(., Covariate = "ANY", Category = "ANY"))) |>
  mutate(n_Covariate = n(),
         .by = Covariate) |>
  mutate(Complex = case_when(Relation %in% c("N", "I", "B") ~ 1,
                               .default = 0)) |>
  summarise(PctComplex = (sum(Complex)/mean(n_Covariate))*100,
            .by = c(Covariate, Category)) |>
  mutate(Complex = TRUE) %>%
  rbind(mutate(., PctComplex = 100-PctComplex, Complex = FALSE)) |>
  mutate(Covariate = case_when(Covariate == "Habitat" ~ "Habitat",
                               Covariate == "Climate" ~ "Climate",
                               Covariate == "Spatial" ~ "Spatial",
                               Covariate == "Human" ~ "Human",
                               Covariate == "HYDR" ~ "Water",
                               Covariate == "BIOT" ~ "Biotic",
                               Covariate == "TOPO" ~ "Topography",
                               Covariate == "PHEN" ~ "Timing",
                               Covariate == "SEAS" ~ "Primary",
                               Covariate == "SURV" ~ "Secondary",
                               Covariate == "OBSE" ~ "Observation",
                               Covariate == "SITE" ~ "Site",
                               Covariate == "SPEC" ~ "Species",
                               Covariate == "ANY" ~ "Any",
                               Covariate == "E" ~ "Any\nEnv.",
                               Covariate == "S" ~ "Any\nStruct."),
         .default = Covariate) |>
  filter(!is.na(Covariate)) |>
  mutate(Covariate = fct(Covariate,
                         levels = c("Any", "Any\nEnv.",
                                    "Habitat", "Spatial", "Climate",
                                    "Timing", "Human", "Topography",
                                    "Biotic", "Water",
                                    "Any\nStruct.", "Primary", "Observation",
                                    "Secondary", "Site", "Species"))) |>
  mutate(Category = case_when(Category == "ANY" ~ "Any",
                              Category == "E" ~ "Environmental",
                              Category == "S" ~ "Structural")) |>

  ggplot(aes(x = Covariate, y = PctComplex, fill = Complex)) +
  geom_col(colour = "white", 
           linewidth = 1, width = 1) +
  scale_fill_manual(limits = c(TRUE, FALSE), 
                    values = c("plum3", "gray50")) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(Category), scales = "free_x", space = "free_x") +
  labs(y = "Percent nonlinear\nor interactive", x = "") +
  theme(legend.position = "none",
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(family = "Helvetica", 
                                   angle = 90, vjust = 0.5, hjust = 1),
        strip.text.x = element_blank(), strip.background = element_blank())
covariateComplex
```

#### Combine

```{r Covariate_Patchwork}
covariatePatchwork <- covariateFreq / covariateDyn / covariateObs / covariateComplex +
  plot_layout(axes = "collect") &
  theme(axis.text = element_text(family = "Helvetica"),
        axis.title = element_text(family = "Helvetica"))
covariatePatchwork

ggsave(plot = covariatePatchwork, "Figures/CovariateSummary.png",
```

@@

```{r Objective proportions}
stat_objectiveProportions <- sheet_Objectives |>
  mutate(TotalArticles = length(unique(`Review ID`))) |>
  pivot_longer(c("Demonstrate methods?", "Estimate trends?", 
                 "Identify drivers?", "Test relationships?",
                 "Predict spatially?", "Predict temporally?"),
               names_to = "Objective", values_to = "ObjectiveIncluded") |>
  filter(ObjectiveIncluded == "YES") |>
  summarise(ObjArticles = n(), 
            .by = c(Objective, TotalArticles)) |>
  mutate(PropArticles = ObjArticles/TotalArticles)
stat_objectiveProportions

# Obj per study
sheet_Objectives |>
  mutate(TotalArticles = length(unique(`Review ID`))) |>
  pivot_longer(c("Demonstrate methods?", "Estimate trends?", 
                 "Identify drivers?", "Test relationships?",
                 "Predict spatially?", "Predict temporally?"),
               names_to = "Objective", values_to = "ObjectiveIncluded") |>
  filter(ObjectiveIncluded == "YES") |>
  summarise(nObj = n(),
            .by = c(`Review ID`, TotalArticles)) |>
  summarise(nFreq = n(),
            .by = c(nObj))
```

@@

## Time Series

#### Number of sites

```{r TimeSeries_Sites}
timeSeries_Sites <- sheet_DataCollection |>
  filter(`Site quantity` != "Unclear") |>
  mutate(`Site quantity` = as.numeric(`Site quantity`)) |>
  select(`Review ID`, SiteNumber = `Site quantity`) |>
  left_join(select(sheet_ArticleData, c(`Review ID`, `Year Strata`, Year)), 
            by = "Review ID") |>

  ggplot(aes(x = `Year Strata`, y = SiteNumber)) +
  geom_boxplot(fill = "goldenrod", colour = "gray40", staplewidth = 0.4,
               linewidth = 0.2, , outlier.size = 0.2) +
  scale_y_log10() +
  labs(x = "", y = "Number of sites\n(log scale)",
       title = "Site quantity") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), 
        axis.title = element_text(size = 8, family = "Helvetica"), 
        axis.text = element_text(size = 8, family = "Helvetica"))
timeSeries_Sites
```

#### Number of covariates considered

```{r TimeSeries_Covariates}
timeSeries_Covariates <- sheet_Covariates |>
  separate_longer_delim(`Model ID`, delim = ",") |>
  separate_longer_delim(Covariates, delim = ", ") |>
  mutate(studyID = paste(`Review ID`, `Model ID`, sep = "_")) |>
  # Get number of core parameters
  filter(`Parameter type` %in% c("Initial occupancy", "Occupancy",
                                 "Colonisation", "Extinction_Persistence",
                                 "Detection")) |>
  mutate(n_params = length(unique(`Parameter type`)), .by = studyID) |>
  # Remove 'none' covariates
  filter(Covariates != "None") |>
  # Calculate number of covariates per parameter
  summarise(CovsPerParam = n()/(max(n_params)), .by = c(studyID, `Review ID`)) |>
  # Models with 0 covariates equate to NA, convert these to 0s
  mutate(CovsPerParam = case_when(is.na(CovsPerParam) ~ 0,
                                  .default = CovsPerParam)) |>
  left_join(select(sheet_ArticleData, c(`Review ID`, `Year Strata`, Year)), 
            by = "Review ID") |>

  ggplot(aes(x = `Year Strata`, y = CovsPerParam)) +
  geom_boxplot(fill = "goldenrod", colour = "gray40", staplewidth = 0.4,
               linewidth = 0.2, outlier.size = 0.2) +
  scale_y_log10() +
  labs(x = "", y = "Covariates\nper parameter",
       title = "Covariates considered") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), 
        axis.title = element_text(size = 8, family = "Helvetica"), 
        axis.text = element_text(size = 8, family = "Helvetica"))
timeSeries_Covariates
```

#### Proportion of selection methods (Frequentist)

```{r TimeSeries_Selection}
timeSeries_Selection <- sheet_Modelling |>
  separate_longer_delim(`Model ID`, delim = ",") |>
  mutate(studyID = paste(`Review ID`, `Model ID`, sep = "_")) |>
  filter(`Model class` == "Frequentist") |>
  separate_longer_delim(`Selection method`, ", ") |>
  mutate(Select = 1) |>
  select(`Selection method`, `Review ID`, studyID, Select) |>
  left_join(select(sheet_ArticleData, c(`Review ID`, `Year Strata`, Year)), 
            by = "Review ID") |>
  mutate(n_strata = length(unique(studyID)),
         .by = `Year Strata`) |>
  summarise(sumMethod = sum(Select),
            .by = c(`Selection method`, `Year Strata`, n_strata)) |>
  mutate(PropMethod = (sumMethod/n_strata)*100) |>
  mutate(`Year Strata` = fct(`Year Strata`)) |>
  filter(`Selection method` %in% c("A priori", "Sequential", 
                                   "Candidate set", "Simple")) |>
  complete(`Year Strata`, `Selection method`, fill = list(PropMethod = 0)) |>

  ggplot(aes(x = `Year Strata`, y = PropMethod,
             colour = `Selection method`, group = `Selection method`)) +
  geom_line() +
  geom_point(size = 1) +
  labs(x = "", y = "Percentage of models",
       title = "Selection method (MLE*)") +
  scale_colour_manual("",
                      limits = c("A priori", "Candidate set", 
                                 "Sequential", "Simple"),
                      labels = c("A priori", "Candidate set",
                                 "Sequential", "Simple precursors"),
                      values = c("aquamarine4", "cadetblue3",
                                 "plum4", "goldenrod")) +
  guides(colour = guide_legend(nrow = 2,
                               override.aes = list(size = 0.5,
                                                   linewidth = 0.5))) +
  theme(legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), 
        axis.title = element_text(size = 8, family = "Helvetica"), 
        axis.text = element_text(size = 8, family = "Helvetica"),
        legend.title = element_text(size = 8, family = "Helvetica"), 
        legend.text = element_text(size = 8, family = "Helvetica"))
timeSeries_Selection
```

#### Goodness-of-fit and validation

```{r TimeSeries_Evaluation}
timeSeries_Validation <- sheet_Modelling |>
  separate_longer_delim(`Model ID`, delim = ",") |>
  mutate(studyID = paste(`Review ID`, `Model ID`, sep = "_")) |>
  separate_longer_delim(`Performance evaluation`, ", ") |>
  mutate(`Performance evaluation` = case_when(`Performance evaluation` %in% c("GOF", "BayesianP") ~ "GOF",
                                              `Performance evaluation` %in% c("Validation_In", "Validation_Out") ~ "Validation",
                                              .default = "None")) |>
  select(`Performance evaluation`, `Review ID`, studyID) |>
  left_join(select(sheet_ArticleData, c(`Review ID`, `Year Strata`, Year)), 
            by = "Review ID") |>
  mutate(n_strata = length(unique(studyID)),
         .by = `Year Strata`) |>
  summarise(sumEval = n(),
            .by = c(`Performance evaluation`, `Year Strata`, n_strata)) |>
  mutate(PropEval = (sumEval/n_strata)*100) |>
  complete(`Year Strata`, `Performance evaluation`, fill = list(PropEval = 0)) |>
  mutate(`Year Strata` = fct(`Year Strata`)) |>
  filter(`Performance evaluation` != "None") |>

  ggplot(aes(x = `Year Strata`, y = PropEval, 
             colour = `Performance evaluation`, group = `Performance evaluation`)) +
  geom_line() +
  geom_point(size = 1) +
  scale_colour_manual("", 
                      limits = c("GOF", "Validation"),
                      labels = c("Goodness-of-fit", "Validation"),
                      values = c("aquamarine4", "goldenrod")) +
  labs(x = "", y = "Percentage of models",
       title = "Model assessment") +
  guides(colour = guide_legend(override.aes = list(size = 0.5,
                                                   linewidth = 0.5))) +
  theme(legend.position = "top", 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), 
        axis.title = element_text(size = 8, family = "Helvetica"), 
        axis.text = element_text(size = 8, family = "Helvetica"),
        legend.title = element_text(size = 8, family = "Helvetica"),
        legend.text = element_text(size = 8, family = "Helvetica"))
timeSeries_Validation
```

#### Framework

```{r TimeSeries_Framework}
timeSeries_Framework <- sheet_Modelling |>
  separate_longer_delim(`Model ID`, delim = ",") |>
  mutate(studyID = paste(`Review ID`, `Model ID`, sep = "_")) |>
  select(`Model class`, `Review ID`, studyID) |>
  left_join(select(sheet_ArticleData, c(`Review ID`, `Year Strata`, Year)), 
            by = "Review ID") |>
  mutate(n_strata = length(unique(studyID)),
         .by = `Year Strata`) |>
  summarise(sumMethod = n(),
            .by = c(`Model class`, `Year Strata`, n_strata)) |>
  mutate(PropMethod = (sumMethod/n_strata)*100) |>
  mutate(`Year Strata` = fct(`Year Strata`)) |>
  complete(`Year Strata`, `Model class`, fill = list(PropMethod = 0)) |>
  filter(`Model class` != "Other") |>

  ggplot(aes(x = `Year Strata`, y = PropMethod,
             colour = `Model class`, group = `Model class`)) +
  geom_line() +
  geom_point(size = 1) + 
  scale_colour_manual("", 
                      limits = c("Frequentist", "Bayesian"),
                      labels = c("Maximum Likelihood", "Bayesian"),
                      values = c("aquamarine4", "goldenrod")) +
  guides(colour = guide_legend(nrow = 2, 
                               override.aes = list(size = 0.2,
                                                   linewidth = 0.2))) +
  labs(x = "", y = "Percentage of models", 
       title = "Model framework") +
  ylim(0,100) +
  theme(legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(), 
        axis.title = element_text(size = 8, family = "Helvetica"), 
        axis.text = element_text(size = 8, family = "Helvetica"),
        legend.title = element_text(size = 8, family = "Helvetica"),
        legend.text = element_text(size = 8, family = "Helvetica"))
timeSeries_Framework
```

#### Combine plots

```{r TimeSeries_Combined}
timeSeries_Combined <- timeSeries_Sites + timeSeries_Covariates + timeSeries_Framework + timeSeries_Selection + free(timeSeries_Validation, "space", "l") +
  plot_layout(design = "1122
                        3344
                        #55#") +
  plot_annotation(tag_levels = "A") & 
  theme(legend.position = "top",
        axis.text.x = element_text(size = 6, family = "Helvetica", 
                                 angle = 45, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 6, family = "Helvetica"),
        axis.title = element_text(size = 6, family = "Helvetica"),
        legend.text = element_text(size = 6, family = "Helvetica"),
        legend.title = element_text(size = 6, family = "Helvetica"),
        legend.key.size = unit(.2, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(0,-10,-10,-10),
        plot.tag = element_text(size = 8, family = "Helvetica"),
        plot.title = element_text(size = 6, family = "Helvetica",
                                  hjust = 0.5))
timeSeries_Combined

ggsave(plot = timeSeries_Combined, "Figures/TimeSeries.png",
       width = 11, height = 14, units = "cm")
```
